### 2021-06-20 draftReport.Rmd
default files

---
title: "CYO Project Report - Brazilian Ecommerce"
author: "Asham Vohra"
date: "6/21/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Clustering:

Some customers didn't write a review. But why are they happy or mad?

Sales Prediction:

With purchase date information you'll be able to predict future sales.

Delivery Performance:

You will also be able to work through delivery performance and find ways to optimize delivery times.

Product Quality: (Let's see if we can find this)

Enjoy yourself discovering the products categories that are more prone to customer insatisfaction.

Attention

- An order might have multiple items.
- Each item might be fulfilled by a distinct seller.
- All text identifying stores and partners where replaced by the names of Game of Thrones great houses.

olist_customers_dataset.csv
This dataset has information about the customer and its location. Use it to identify unique customers in the orders dataset and to find the orders delivery location.
At our system each order is assigned to a unique customerid. This means that the same customer will get different ids for different orders. The purpose of having a customerunique_id on the dataset is to allow you to identify customers that made repurchases at the store. Otherwise you would find that each order had a different customer associated with.

olist_order_reviews_dataset.csv
This dataset includes data about the reviews made by the customers.

After a customer purchases the product from Olist Store a seller gets notified to fulfill that order. Once the customer receives the product, or the estimated delivery date is due, the customer gets a satisfaction survey by email where he can give a note for the purchase experience and write down some comments.

olist_orders_dataset.csv
This is the core dataset. From each order you might find all other information.

olist_products_dataset.csv
Products Dataset
This dataset includes data about the products sold by Olist.

# Overview
Data set: https://www.kaggle.com/olistbr/brazilian-ecommerce

# Analysis
# Results
# Conclusion

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.