### 2021-06-27 draftReport.Rmd
added transformations for geo data

---
title: "CYO Project Report - Brazilian Ecommerce"
author: "Asham Vohra"
date: "6/21/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Clustering:

Some customers didn't write a review. But why are they happy or mad?

Sales Prediction:

With purchase date information you'll be able to predict future sales.

Delivery Performance:

You will also be able to work through delivery performance and find ways to optimize delivery times.

Product Quality: (Let's see if we can find this)

Enjoy yourself discovering the products categories that are more prone to customer insatisfaction.

Attention

- An order might have multiple items.
- Each item might be fulfilled by a distinct seller.
- All text identifying stores and partners where replaced by the names of Game of Thrones great houses.

olist_customers_dataset.csv
This dataset has information about the customer and its location. Use it to identify unique customers in the orders dataset and to find the orders delivery location.
At our system each order is assigned to a unique customerid. This means that the same customer will get different ids for different orders. The purpose of having a customerunique_id on the dataset is to allow you to identify customers that made repurchases at the store. Otherwise you would find that each order had a different customer associated with.

olist_order_reviews_dataset.csv
This dataset includes data about the reviews made by the customers.

After a customer purchases the product from Olist Store a seller gets notified to fulfill that order. Once the customer receives the product, or the estimated delivery date is due, the customer gets a satisfaction survey by email where he can give a note for the purchase experience and write down some comments.

olist_orders_dataset.csv
This is the core dataset. From each order you might find all other information.

olist_products_dataset.csv
Products Dataset
This dataset includes data about the products sold by Olist.

# Overview
Data set: https://www.kaggle.com/olistbr/brazilian-ecommerce

# Analysis

Product Quality: (Let's see if we can find this)

Trying to identify the products categories that are more prone to customer insatisfaction. Some things come to mind
- categories with more poorly rated products? data set--> order_reviews_dataset
- categories which tend to be delivered late? data set-->
- expensive products maybe? do not get value for money?
- bulky products maybe?
- some categories maybe limited to few sellers and they may not be selling good products?

Likely classification problem that you select category and given paramters should it come undder customer satisfaction or not.
Anything less than 3 is unsatisfied
3 or more is satisfied

but for same product you may have some people satisfied and some as not. 
( I think that's where knn and related algorithms were introduced, to find mean of the points)

Would be good clustering project but may not be ideal as it based on unsupervised learning and its  coverage was limited.

Maybe it would be wiser to look at sales or delivery times problems.

Clustering:

So given order of an item, delivery time, product etc  can we try to predict customer rating. But will be so miliar to movielens project can become boring

Delivery Performance:

You will also be able to work through delivery performance and find ways to optimize delivery times.

Let's predict delivery time/date
-- delivery distance
-- location of customer as some can be remotely connected and need special delivery merchanism. This is independent of delivery distance.
-- location of seller as shipping from some locations maybe difficult. This is independent of delivery distance.
-- size of product? 
-- price of product (expensive nature of product as it could mean extra care or it maybe fragile?)
-- purchase time....if in night, delivery time maybe longer
-- order approved at time-> if payment approval came multiple hours after purchase time, this maybe longer
-- product category -- products of same category may show similar behavior.
-- or specific product - some products may have higher delivery time.
-- deliveries maybe impacted due to sales, festival season.

we should consider if we need time stamp or we can take purchase time as 0 and calculated approved at from it. though we may lose granularity in terms of time as truck may come everyday at 4. so after 4 you can ship , before 4 you cant

estimated delivery date is present in ---> order_estimated_delivery_date
We can calculate RMSE using that and it will be our benchmark

We can see if we improved than that or not. It will give us indication of how close are we to the production model RMSE.

```{r Install or load packages}
# TODO use conditional loading or installation of library
library(dplyr)
library(lubridate)
library(tidyverse)
install.packages("geosphere")
library(geosphere)
```

```{r Basic analysis}

products_dataset <- read.csv(file.path('data','olist_products_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% mutate(product_category_name = as.factor(product_category_name))

products_dataset %>% head()
products_dataset %>% str()

order_reviews_dataset <- read.csv(file.path('data','olist_order_reviews_dataset.csv'),header=TRUE,stringsAsFactors = FALSE)

order_reviews_dataset %>% head()
order_reviews_dataset %>% str()

# customers data set  with zip code prefix and can be used to uniquely identify a customer
# customer_id ----> key to the orders dataset. Each order has a unique customer_id.
# customer_zip_code_prefix --> first five digits of customer zip code
customers_dataset <- read.csv(file.path('data','olist_customers_dataset.csv'),header=TRUE) %>% mutate(customer_zip_code_prefix=as.factor(customer_zip_code_prefix),customer_id=as.character(customer_id ))

customers_dataset %>% head()
customers_dataset %>% str()

# geo location details
# in brazil while 5 digits zip code is good enough to identify state, it may not be correct to identify city
# it can have geolocation_zip_code_prefix shared across multiple places
# refer: https://codigo-postal.org/en-us/brazil/
# The Postal Code Address (CEP) is an eight-digit number set that has by main objective to guide and accelerate the distribution of objects sent through the national mail in Brazil. The current structure of the Brazilian postal code is 8 (eight) digits, divided into two parts:

# first part composed of 5 digits where each digit represents the 1) region, 2) sub-region, 3) sector, 4) subsector, 5) subsector divider;
# the second part is composed of 3 digits, separated by a hyphen from the first part and represents the Distribution Identifiers.
# You can navigate through states, then cities and finally neighborhoods and streets until you find the corresponding postal code.

geolocation_dataset <- read.csv(file.path('data','olist_geolocation_dataset.csv'),header=TRUE,stringsAsFactors = FALSE)

geolocation_dataset %>% head()
geolocation_dataset %>% str()

# Important: many geolocation_city have same name but different ascent component. Example: niteroi and niterói
# intially we can see 
geolocation_dataset %>% dplyr::count(geolocation_zip_code_prefix,geolocation_city,geolocation_state) %>% nrow()

# post transformation to resolve any Accent or special characters from the city name
geolocation_dataset <- geolocation_dataset %>% mutate(geolocation_city=stri_trans_general(geolocation_city,"Any-ASCII"))

geolocation_dataset %>% dplyr::count(geolocation_zip_code_prefix,geolocation_city,geolocation_state) %>% nrow()

# as each geolocation_zip_code_prefix can have many data points, we will limit these to one by taking mean for lat, long location
geolocation_dataset <- geolocation_dataset %>% group_by(geolocation_zip_code_prefix,geolocation_city,geolocation_state) %>% summarise(geolocation_lat=mean(geolocation_lat),geolocation_lng=mean(geolocation_lng)) %>% ungroup()

# TODO review it
#%>% mutate(geolocation_state=as.factor(geolocation_state)) 
# TODO review if geolocation_zip_code_prefix and geolocation_city can have multiple states with them.

# remove whitespace before and after reduces one city
# still issues are there and can be seen by
zip_codes_of_interest <- geolocation_dataset %>% mutate(geolocation_city=str_trim(geolocation_city)) %>% dplyr::count(geolocation_zip_code_prefix) %>% filter(n>1)  %>% pull(geolocation_zip_code_prefix)

geolocation_dataset %>% filter(geolocation_zip_code_prefix %in% zip_codes_of_interest) %>% view()
# to filter out duplicates
# we need not worry about city or state, we can pick any as long as lat, lng values are similar

# seller data set with zip code prefix
# seller_zip_code_prefix --> first 5 digits of seller zip code
sellers_dataset <- read.csv(file.path('data','olist_sellers_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) 
# TODO review if we need it
# %>% mutate(seller_zip_code_prefix=as.factor(seller_zip_code_prefix),seller_id=as.character(seller_id ))

sellers_dataset %>% head()
sellers_dataset %>% str()

# Important: many geolocation_city have same name but different ascent component. Example: niteroi and niterói
# These all need to be handled in data wrangling

sellers_dataset %>% inner_join(geolocation_dataset,by=c('seller_zip_code_prefix'='geolocation_zip_code_prefix')) %>% str()

# TODO there seems to be an issue
sellers_dataset %>% inner_join(geolocation_dataset,by=c('seller_zip_code_prefix'='geolocation_zip_code_prefix')) %>% dplyr::count(seller_id) %>% slice_max(order_by=n,n=5)

# geolocation_dataset seems to have 4 entries for same zip code . example 13454

# we tried another approach by joining over all zip codee, city and state but it led to no match
sellers_dataset %>% mutate(seller_city=stri_trans_general(seller_city,"Any-ASCII")) %>% inner_join(geolocation_dataset,by=c('seller_zip_code_prefix'='geolocation_zip_code_prefix','seller_city'='geolocation_city','seller_state'='geolocation_state')) %>% str()
# TODO -- if we analyse we can see city and state for seller data differ in both datasets. so best to stick to zip code. city and state are prone to errors.
# we used anti join to identify ones not matching.

sellers_dataset %>% anti_join(geolocation_dataset,by=c('seller_zip_code_prefix'='geolocation_zip_code_prefix'))

# order data set with various delivery related times
# note each customer_id is unique per order
# order_purchase_timestamp -> Shows the purchase timestamp.
# order_approved_at/payment_approved_at -> Shows the payment approval timestamp.
# order_delivered_carrier_date/handed_over_to_carrier_date -> Shows the order posting timestamp. When it was handled to the logistic partner.
# order_delivered_customer_date --> Shows the actual order delivery date to the customer.
# order_estimated_delivery_date --> Shows the estimated delivery date that was informed to customer at the purchase moment.
orders_dataset <- read.csv(file.path('data','olist_orders_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% rename(payment_approved_at=order_approved_at,handed_over_to_carrier_date=order_delivered_carrier_date)  %>% mutate(order_status=as.factor(order_status),order_purchase_timestamp=as_datetime(order_purchase_timestamp),payment_approved_at=as_datetime(payment_approved_at),handed_over_to_carrier_date=as_datetime(handed_over_to_carrier_date),order_delivered_customer_date=as_datetime(order_delivered_customer_date),order_estimated_delivery_date=as_datetime(order_estimated_delivery_date))

orders_dataset <- orders_dataset %>% filter(order_status=='delivered') %>% filter( !is.na(order_estimated_delivery_date ) & !is.na(order_delivered_customer_date ) )

orders_dataset %>% head()
orders_dataset %>% str()

# current RMSE for predicted delivery time given by order_estimated_delivery_date vs actual outcome give by order_delivered_customer_date
# However, note that order_estimated_delivery_date has no timestamp, so we will remove timestamp component, round date using floor_date, which means any delivery done during the day say at 2017-10-10 21:25:13 is treated to be done on 2017-10-10

# first we convert it to desired date
# then we covert date to numeric i.e. to convert it to seconds
orders_dataset %>% mutate(order_delivered_customer_date=floor_date(order_delivered_customer_date)) %>% mutate(order_delivered_customer_date=as.numeric(order_delivered_customer_date),order_estimated_delivery_date=as.numeric(order_estimated_delivery_date))  %>% head()

# then we calculate time_taken_from_order_to_delivery, estimated_time_from_order_to_delivery in seconds

orders_dataset %>% head() %>% mutate(order_delivered_customer_date=floor_date(order_delivered_customer_date),order_purchase_timestamp=as.numeric(order_purchase_timestamp)) %>%  mutate(time_taken_from_order_to_delivery=as.numeric(order_delivered_customer_date)-order_purchase_timestamp, estimated_time_from_order_to_delivery = as.numeric(order_estimated_delivery_date) -order_purchase_timestamp) %>% str()

orders_dataset %>% mutate(order_delivered_customer_date=floor_date(order_delivered_customer_date,unit="day"),order_purchase_timestamp=as.numeric(order_purchase_timestamp)) %>%  mutate(time_taken_from_order_to_delivery=as.numeric(order_delivered_customer_date)-order_purchase_timestamp, estimated_time_from_order_to_delivery = as.numeric(order_estimated_delivery_date) -order_purchase_timestamp,diff=estimated_time_from_order_to_delivery-time_taken_from_order_to_delivery)

# if we look at graph we can  see there is huge difference between estimate data and actual delivery date
# are these different readings or mode of transport is different? 
p1 <- orders_dataset %>% mutate(order_delivered_customer_date=floor_date(order_delivered_customer_date,unit="day"),order_purchase_timestamp=as.numeric(order_purchase_timestamp)) %>%  mutate(time_taken_from_order_to_delivery=as.numeric(order_delivered_customer_date)-order_purchase_timestamp, estimated_time_from_order_to_delivery = as.numeric(order_estimated_delivery_date) -order_purchase_timestamp,diff=estimated_time_from_order_to_delivery-time_taken_from_order_to_delivery) %>% ggplot(aes(diff)) + geom_histogram(binwidth=86400*15)

library(plotly)
ggplotly(p1)

# orders data with item related seller info
# order_item_id --->sequential number identifying number of items included in the same order.
# so if there are three items they will be nuber 1 , 2, 3 for the given order
# seller_limit_date ---> Shows the seller shipping limit date for handling the order over to the logistic partner.
# freight_value --> item freight value item (if an order has more than one item the freight value is splitted between items)
order_items_dataset <- read.csv(file.path('data','olist_order_items_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% mutate(order_id=as.factor(order_id),product_id=as.factor(product_id),seller_id=as.factor(seller_id))

order_items_dataset %>% str()
order_items_dataset %>% head()

# we can combine orders delivery information (orders_dataset) with orders items and seller related info (order_items_dataset)
orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% head()

# let's combine with seller details
orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% head()

# let's combine with unique customer and its location dataset
orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% select(-customer_id) %>% head()

orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% select(-customer_id) %>% head() %>% mutate(distance_between_two_points=round(distGeo(c(-80,40),c(-81,41))/1000,2))

# we will use distGeo from geosphere package to compute distance in km between seller and customer. considered to be highly accurate. if it takes more time, we can consider using distVincentyEllipsoid or distHaversine as we are fine with accuracy of a km.
orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% select(-customer_id) %>% head(2) %>% mutate(seller_zip_code_prefix=as.numeric(seller_zip_code_prefix)) %>% select('seller_zip_code_prefix') %>% inner_join(geolocation_dataset,by=c('seller_zip_code_prefix'='geolocation_zip_code_prefix')) %>% rename(geolocation_lat_seller=geolocation_lat,geolocation_lng_seller=geolocation_lng)

orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% select(-customer_id) %>% head(5) %>% mutate(seller_zip_code_prefix=as.numeric(seller_zip_code_prefix)) %>% mutate(customer_zip_code_prefix=as.numeric(customer_zip_code_prefix)) %>% inner_join(geolocation_dataset,by=c('seller_zip_code_prefix'='geolocation_zip_code_prefix')) %>% rename(geolocation_lat_seller=geolocation_lat,geolocation_lng_seller=geolocation_lng) %>% select(customer_zip_code_prefix) %>% inner_join(geolocation_dataset,by=c('customer_zip_code_prefix'='geolocation_zip_code_prefix')) %>% rename(geolocation_lat_customer=geolocation_lat,geolocation_lng_customer=geolocation_lng)

```

# Results
# Conclusion

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.